A Documentary History of \\ the Nock Combinator Calculus

### 2025-08-13 mss.tex
Fix various

Nock is a combinator calculus which serves as the computational specification layer for the Urbit and Nockchain/NockApp systems.  It is a hyper-\textsc{risc} instruction set architecture (ISA) intended for execution by a virtual machine (but see \citet{Mopfel2025}, pp.~XX–XX herein).  Nock's simplicity and unity of expression make it amenable to proof-based reasoning and guarantees of correctness.  Its Lisp-like nature surfaces the ability to introspect on the code itself, a property which higher-level languages compiling to it can exploit.  Yet for all this, Nock was not born from a purely mathematical approach, but found its roots in practical systems engineering.

@@

Extensive commentary on the operators is provided.  Rightwards grouping of tuple expressions has already been introduced.  Extension of the language is summarily ruled out.\footnote{Compare Ax and Conk, pp.~XX–XX herein.}  Data are conceived of as Unix-like byte streams; details of parsing and lexing are considered.  Terms (the ancestor of nouns) include a \texttt{NULL}-like ``foo'' type \texttt{\textasciitilde} distinguishable by value rather than structure.  \textsc{ascii} is built in as numeric codes, similar to G\"{o}del numbering.

@@

Source:  \citet{Yarvin2006}

@@

Axiomatic operator \texttt{*} tar\footnote{We refer to Nock axiomatic operators via their modern aural \textsc{ascii} pronunciations.  While these evolved over time (to wit, \texttt{\textasciicircum} ``hat'' became ``ket''), to attempt to synchronize pronunciation with the era of a Nock release is a fool's errand.} is identified as a \texttt{GOTO}.\footnote{One can see the influence of this version's naming scheme on \patp{barpub-tarber}'s Ax, pp.~189–226 herein.}

@@

Source:  \citet{Nock13K}

@@

Source:  \citet{Nock12K}

@@

Source:  \citet{Nock11K}

@@

Source:  \citet{Nock10K}

@@

Source:  \citet{Nock9K}

@@

Source:  \citet{Nock8K}

@@

Source:  \citet{Nock7K}

@@

Source:  \citet{Nock6K}

@@

Source:  \citet{Nock5K}

@@

Source:  \citet{Nock4K}

@@

While deviations from the trunk line of the Nock family have been proposed at various points,\footnote{Notably, Ax (see \patp{barpub-tarber}, pp.~XX–XX in this volume), \textsc{skew}, and \textsc{plan} (see \texttt{@sol\_plunder}, pp.~227–XXX in this volume).} Nock itself has remained the definitional substrate of Urbit since its inception.  It has also been adopted as the primary \textsc{isa} of Nockchain and the NockApp ecosystem.

>>>>

\sloppy
Nock is a combinator calculus which serves as the computational specification layer for the Urbit and Nockchain/NockApp systems.  It is a hyper-\textsc{risc} instruction set architecture (ISA) intended for execution by a virtual machine (but see \patp{mopfel-winrux}, pp.~99–130 in this volume).  Nock's simplicity and unity of expression make it amenable to proof-based reasoning and guarantees of correctness.  Its Lisp-like nature surfaces the ability to introspect on the code itself, a property which higher-level languages compiling to it can exploit.  Yet for all this, Nock was not born from a purely mathematical approach, but found its roots in practical systems engineering.

@@

Extensive commentary on the operators is provided.  Rightwards grouping of tuple expressions has already been introduced.  Extension of the language is summarily ruled out.\footnote{Compare Ax and Conk by \patp{barpub-tarber}, pp.~191–227 in this volume.}  Data are conceived of as Unix-like byte streams; details of parsing and lexing are considered.  Terms (the ancestor of nouns) include a \texttt{NULL}-like ``foo'' type \texttt{\textasciitilde} distinguishable by value rather than structure.  \textsc{ascii} is built in as numeric codes, similar to G\"{o}del numbering.

@@

Source:  \citet{Yarvin2006}.

@@

Axiomatic operator \texttt{*} tar\footnote{We refer to Nock axiomatic operators via their modern aural \textsc{ascii} pronunciations.  While these evolved over time (to wit, \texttt{\textasciicircum} ``hat'' became ``ket''), to attempt to synchronize pronunciation with the era of a Nock release is a fool's errand.} is identified as a \texttt{GOTO}.\footnote{One can see the influence of this version's naming scheme on \patp{barpub-tarber}'s Ax, pp.~191–227 in this volume.}

@@

Source:  \patp{sorreg-namtyv} (2008d). %\citet{Nock13K}

@@

Source:  \patp{sorreg-namtyv} (2008c). %\citet{Nock12K}

@@

Source:  \patp{sorreg-namtyv} (2008b). %\citet{Nock11K}

@@

Source:  \patp{sorreg-namtyv} (2008a). %\citet{Nock10K}

@@

Source:  \patp{sorreg-namtyv} (2010c). %\citet{Nock9K}

@@

Source:  \patp{sorreg-namtyv} (2010b). %\citet{Nock8K}

@@

Source:  \patp{sorreg-namtyv} (2010a). %\citet{Nock7K}

@@

Source:  \patp{sorreg-namtyv} (2011). %\citet{Nock6K}

@@

Source:  \patp{sorreg-namtyv} (2012). %\citet{Nock5K}

@@

Source:  \patp{sorreg-namtyv} (2018). %\citet{Nock4K}

@@

While deviations from the trunk line of the Nock family have been proposed at various points,\footnote{Notably, Ax (see \patp{barpub-tarber}'s Ax, pp.~191–227 in this volume), \textsc{skew}, and \textsc{plan} (see \texttt{@sol\_plunder}, pp.~227–XXX in this volume).} Nock itself has remained the definitional substrate of Urbit since its inception.  It has also been adopted as the primary \textsc{isa} of Nockchain and the NockApp ecosystem.
